\documentclass[basic, header]{nosvagor-notes}
\usepackage{nosvagor-math}

\colorlet{title-color}{red}
\newcommand{\theTitle}{%
  \href{https://github.com/nosvagor/notes}%
  {Homework 3}%
}

\newcommand{\userName}{Cullyn Newman}
\newcommand{\class}{STAT: 361}
\newcommand{\institution}{Portland State}

\begin{document}

\begin{enumerate}[itemsep=4em]

  \item (2.85) The probability that a doctor correctly diagnoses a particular
    illness is \(0.7 = P(C)\). Given that the doctor makes an incorrect diagnosis, the
    probability that the patient files a lawsuit is \(0.9 = P(L \mid C')\) . What is the
    probability that the doctor makes an incorrect diagnosis and the patient
    sues \(P(S)\)?
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      P(S) = P(C' \cap L ) = P(L\mid C')P(C') = (0.9)(0.3) = \boxed{0.27}
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \item (2.89) A town has two fire engines operating independently. The
    probability that a specific engine is available when needed is 0.96.

    \begin{enumerate}

      \item What is the probability that neither is available when needed?
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          P(U) = P(A' \cap B') = (0.04)(0.04) = \boxed{0.0016}
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

      \item What is the probability that a fire engine is available when
        needed?
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          P(U') = 1 - 0.0016 = \boxed{0.9984}
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \end{enumerate}

  \item (2.109) A large industrial firm uses three local motels to provide
    overnight accommodations for its clients. From past experience it is known
    that 20\% of the clients are assigned rooms at the Ramada Inn, 50\% at the
    Sheraton, and 30\% at the Lakeview Motor Lodge. If the plumbing is faulty
    in 5\% of the rooms at the Ramada Inn, in 4\% of the rooms at the Sheraton,
    and in 8\% of the rooms at the Lakeview Motor Lodge, what is the
    probability that
      \begin{align*}
        P(R) = 0.2, P(F_r) = 0.05, P(F_r \cap R) = 0.01 \\
        P(S) = 0.5, P(F_s) = 0.04, P(F_s \cap S) = 0.02 \\
        P(L) = 0.3, P(F_l) = 0.08, P(F_l \cap L) = 0.024
      \end{align*}
    \begin{enumerate}

      \item a client will be assigned a room with faulty plumbing?
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          P(F) = 0.01 + 0.02 + 0.024 = \boxed{0.054}
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

      \item a person with a room having faulty plumbing was assigned
        accommodations at the Lakeview Motor Lodge?
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          P(L \mid F) = \frac{P(F_l \cap L)}{P(F)} = \frac{0.024}{0.54} = \boxed{\frac{4}{9}}
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \end{enumerate}

%%%%%%%%%%%%%
  \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%

  \item (2.113) From a box containing 6 black balls and 4 green balls, 3 balls
    are drawn in succession, each ball being replaced in the box before the
    next draw is made. What is the probability that
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      P(B_i) = \frac{6}{10}, \quad P(G_i) = \frac{4}{10}
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{enumerate}

      \item all 3 are the same color?
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          P(A) = P\left((B_1 \cap B_2 \cap B_3) \cup (G_1 \cap G_2 \cap G_3)\right)
          = 0.6^3 + 0.4^3 = \boxed{0.28}
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

      \item each color is represented?
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          P(A') = 1 - 0.28 = \boxed{0.72}
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \end{enumerate}

  \item (3.1) Classify the following random variables as discrete or
    continuous:

    \begin{enumerate}[label={}, leftmargin=0em]

      \item X: the number of automobile accidents per year in Virginia.
        \begin{itemize}
          \item \textbf{discrete}

        \end{itemize}
      \item Y: the length of time to play 18 holes of golf.
        \begin{itemize}
          \item \textbf{continuous}

        \end{itemize}
      \item M: the amount of milk produced yearly by a particular cow.
        \begin{itemize}
          \item \textbf{continuous}

        \end{itemize}
      \item N: the number of eggs laid each month by a hen.
        \begin{itemize}
          \item \textbf{discrete}

        \end{itemize}
      \item P: the number of building permits issued each month in a certain city.
        \begin{itemize}
          \item \textbf{discrete}

        \end{itemize}
      \item Q: the weight of grain produced per acre.
        \begin{itemize}
          \item \textbf{continuous}

        \end{itemize}

    \end{enumerate}

%%%%%%%%%%%%%
  \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%

  \item (3.3) Let \(W\) be a random variable giving the number of heads minus the
    number of tails in three tosses of a coin. List the elements of the sample
    space \(S\) for the three tosses of the coin and to each sample point assign a
    value \(w\) of \(W\).
    \begin{align*}
      S = \left\{ HHH, HHT, HTH, THH, TTH, THT, TTH, TTT \right\} \\
      W = \left[ 3, 1, 1, 1, -1, -1, -1, -3 \right] \\
      S_i \to W_i
    \end{align*}

  \item (3.35) Suppose it is known from large amounts of historical data that
    \(X\), the number of cars that arrive at a specific intersection during a
    20 second time period, is characterized by the following discrete
    probability function:
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      f(x) = e^{-6}\frac{6^x}{x!}, \forall x \geq 0 \in \Z
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{enumerate}

      \item Find the probability that in a specific 20 second
        time period, more than 8 cars arrive at the intersection.
        \begin{align*}
          P(x \leq 8) = \sum_{x=0}^{8} e^{-6} \frac{6^{x}}{x!} = 0.847 \\
          P(x > 8) = 1 - P(x \leq 8) = \boxed{0.153}
        \end{align*}

      \item Find the probability that only 2 cars arrive.
        \begin{align*}
          f(2) = e^{-6}\frac{6^2}{2!} = \boxed{0.045}
        \end{align*}

    \end{enumerate}


\end{enumerate}


\end{document}
