\documentclass[basic, header]{nosvagor-notes}
\usepackage{nosvagor-math}

\colorlet{title-color}{red}
\newcommand{\theTitle}{%
  \href{https://github.com/nosvagor/notes}%
  {Homework 4}%
}

\newcommand{\userName}{Cullyn Newman}
\newcommand{\class}{STAT: 361}
\newcommand{\institution}{Portland State}

\begin{document}

\begin{enumerate}[leftmargin=1.5em, itemsep=4em]

  \item \textbf{4.35} The random variable X, representing the number of
    errors per 100 lines of software code, has the following probability
    distribution:

    \begin{table}[h]
      \centering
      \begin{tabular}{c|ccccc}
        x &  2 & 3 & 4  & 5 & 6 \\
        \midrule
        \(f(x)\)  & 0.01 & 0.25 &0.4 & 0.3 & 0.04 \\
      \end{tabular}
    \end{table}

    Using Theorem 4.2 on page 121, find the variance of \(X\).

  \item \textbf{4.36} Suppose that the probabilities are \(0.4, 0.3, 0.2,
    ~\text{and}~ 0.1\), respectively, that \(0, 1, 2, ~\text{or}~ 3\) power
    failures will strike a certain subdivision in any given year.

    Find the mean and variance of the random variable \(X\) representing the
    number of power failures striking this subdivision.

  \item \textbf{4.37} A dealer’s profit, in units of \$5000, on a new
    automobile is a random variable \(X\) having the density function given in
    Exercise 4.12 on page 117.

    Find the variance of \(X\).

%%%%%%%%%%%%%
  \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%

  \item \textbf{4.38} The proportion of people who respond to a certain
    mail-order solicitation is a random variable \(X\) having the density
    function given in Exercise 4.14 on page 117.

    Find the variance of \(X\).

  \item \textbf{4.43 (Bonus)} The length of time, in minutes, for an airplane
    to obtain clearance for takeoff at a certain airport is a random variable
    \(Y = 3X - 2\), where \(X\) has the density function
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      f(x) =
      \begin{cases}
        \frac{1}{4}e^{-\frac{x}{4}}, & x > 0 \\
        0, & ~\text{elsewhere}~
      \end{cases}
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    Find the mean and variance of the random variable \(Y\).

  \item \textbf{4.50} For a laboratory assignment, if the equipment is working,
    the density function of the observed outcome \(X\) is
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      f(x) =
      \begin{cases}
        2(1-x), & 0 < x < 1, \\
        O,      & \text{elsewhere}
      \end{cases}
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    Find the variance and standard deviation of \(X\).

%%%%%%%%%%%%%
  \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%

  \item \textbf{4.54} Using Theorem 4.5 and Corollary 4.6, find the mean and
    variance of the random variable \(Z = 5X +3\), where \(X\) has the probability
    distribution of Exercise 4.36 on page 127.

  \item \textbf{4.71 (Bonus)} The length of time \(Y\), in minutes, required
    to generate a human reflex to tear gas has the density function
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      f(y) =
      \begin{cases}
        \frac{1}{4}e^{-\frac{y}{4}}, & 0 \leq y < \infty \\
         0, & \text{elsewhere}
      \end{cases}
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{enumerate}[leftmargin=1.6em]

      \item What is the mean time to reflex?

      \item Find \(E(Y^2)\) and \(\var(Y)\).

    \end{enumerate}

  \item \textbf{4.101} Consider Review Exercise 3.73 on page 108. It
    involved \(Y\), the proportion of impurities in a batch, and the density
    function is given by
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      f(y) =
      \begin{cases}
        10(1-y)^9,			& 0 \leq y < 1, \\
        0, & \text{elsewhere}
      \end{cases}
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{enumerate}[leftmargin=1.6em]

      \item Find the expected percentage of impurities.

      \item Find expected value of proportion of quality material, i.e.,
        \(E(1-Y)\).

    \end{enumerate}

%%%%%%%%%%%%%
  \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%

  \item \textbf{4.62} If \(X\) and \(Y\) are independent random variables with
    variances \(\sigma^2_X = 5 ~\text{and}~ \sigma^2_Y = 3\), find the variance
    of the random variable \(Z = -2X + 4Y - 3\).

  \item \textbf{4.63} Repeat Exercise 4.62 if X and Y are not independent and
    \(\sigma_{XY} = 1\).

  \item Let \(X\) and \(Y\) be random variables with the following information:
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \hspace{-4em} E(X) = 6, \quad E(Y) = -\frac{1}{2}, \quad \var(X) = 4, \quad \var(Y) = 6, \quad \cov(X,Y) = 2
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{enumerate}[leftmargin=1.6em]

      \item Compute \(E(3X - 4Y)\)
      \item Compute \(\var(3X - 4Y)\)
      \item Compute \(E(2X - Y^2)\)

    \end{enumerate}

%%%%%%%%%%%%%
  \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%

  \item Let \(X\) and \(Y\) be independent random variables with the following
    information:
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     E(X) = -1, \quad E(Y) = 4, \quad \var(X) = 6, \quad \var(Y) = 8
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{enumerate}[leftmargin=1.6em]

      \item Compute \(E(9X + 2Y)\)
      \item Compute \(\var(9X + 2Y)\)

    \end{enumerate}

  \item \textbf{6.3} The daily amount of coffee, in liters, dispensed by a
    machine located in an airport lobby is a random variable \(X\) having a
    continuous uniform distribution with \(A = 7 ~\text{and}~ B = 10\).

    Find the probability that on a given day the amount of coffee dispensed by
    this machine will be
    \begin{enumerate}[leftmargin=1.6em]

      \item at most 8.8 liters;

      \item more than 7.4 liters but less than 9.5 liters;

      \item at least 8.5 liters.

    \end{enumerate}

  \item \textbf{6.4} A bus arrives every 10 minutes at a bus stop. It is
    assumed that the waiting time for a particular individual is a random
    variable with a continuous uniform
    \begin{enumerate}[leftmargin=1.6em]

      \item What is the probability that the individual waits more than 7
        minutes?

      \item  What is the probability that the individual waits between 2 and 7
        minutes?

    \end{enumerate}

\end{enumerate}

\end{document}
