\chapter{5 Eigenvalues and Eigenvectors}

\section{5.1 Introduction to Eigenvalues and Eigenvectors}
\begin{itemize}
  \item[]

  \subsection{Problems 39, 40}
  \begin{enumerate}
    \minor{\item[39.]  Challenge problem: Is there a real 2 by 2 matrix (other
      than \(I\)) with \(\bm{A}^3 = I\)? Its eigenvalues must satisfy \(\lambda^3 =
      I\). They can be \(e^{ 2\pi\i /3}\) and \( e^{-2\pi\i i/3}\).

      What trace and determinant would this give? Construct \tbm{A}.
    }
    \begin{align*}
      \det{\bm{A}} = e^{\frac{2\pi\i}{3}} \cdot e^{-\frac{2\pi\i}{3}} = e^0 = 1 \\
      \tr{\bm{A}} = e^{\frac{2\pi\i}{3}} + e^{-\frac{2\pi\i}{3}} = -1 \\
      \text{A rotation matrix satisfies the condition, thus:}~ \\
      \bm{A} =
      \begin{bmatrix}
        \cos \frac{2\pi}{3}  & -\sin \frac{2\pi}{3}  \\
        \sin \frac{2\pi}{3} & \cos \frac{2\pi}{3}
      \end{bmatrix}
    \end{align*}


    \minor{\item[40.] There are six 3 by 3 permutation matrices \tbm{P}.

      \fg{Note: \ulink{s:Problems 14, 18}{Section 1.5, problem 14} lists the 6
        permutation matrices; all of which can represented as the product of
        elementary matrices.}

      \begin{enumerate}
        \item What numbers can be the determinants of \tbm{P}?

        \begin{itemize}\color{foreground}
          \item Multiplication by an elementary matrix at most produces a sign flip,
          thus, the determinants are \(\pm 1\).
        \end{itemize}

        \item What numbers can be pivots?
        \begin{itemize}\color{foreground}
          \item All pivots must be equal to 1 (only nonzero number available).
        \end{itemize}

        \item What numbers can be the trace of \tbm{P}?
        \begin{itemize}\color{foreground}
          \item Ranges from 0--3, depends on how many 1's are present in the diagonal.
        \end{itemize}

        \item What numbers can be eigenvalues of \tbm{P}?
        \begin{itemize}\color{foreground}
          \item \(\lambda^3 - 1 = 0\), thus:
            \begin{itemize}
              \item \(\lambda_1 = 1\)
              \item \(\displaystyle \lambda_2 = -\frac{1}{2} - \frac{i\sqrt{3} }{2}\)
              \item \(\displaystyle \lambda_3 = -\frac{1}{2} + \frac{i\sqrt{3}}{2}\)
            \end{itemize}
        \end{itemize}

      \end{enumerate}
    }
  \end{enumerate}
\end{itemize}

\section{5.2 Diagonalization of a Matrix}
\begin{itemize}
  \item []

  \subsection{Problems 11, 14, 39}
  \begin{enumerate}
    \minor{\item[11.] If the eigenvalues of \tbm{A} are 1, 1, 2, which of the
      following are certain to be true? Give a reason if true or a
      counterexample if false:
      \begin{enumerate}
        \item \tbm{A} is invertible.
        \begin{itemize}\color{foreground}
          \item \true{true}; none of the eigenvalues are zero, meaning the
            determinant is nonzero, which means \tbm{A} is non-degenerate (invertible).
        \end{itemize}

        \item \tbm{A} is diagonalizable.
        \begin{itemize}\color{foreground}
          \item \false{false}; depends on number of linearly independent
            igenvectors.
          \[%%%%%%%%%%
          \bm{A} =
          \begin{bmatrix}
            1 & 1 & 1\\
            0 & 1 & 1 \\
            0 & 0 & 2
          \end{bmatrix}
          \]%%%%%%%%%%
        \end{itemize}

        \item \tbm{A} is not diagonalizable.
        \begin{itemize}\color{foreground}
          \item \false{false}; the following is certainly a diagonal matrix.
          \[%%%%%%%%%%
          \bm{A} =
          \begin{bmatrix}
            1 & 0 & 0\\
            0 & 1 & 0 \\
            0 & 0 & 2
          \end{bmatrix}
          \]%%%%%%%%%%
        \end{itemize}

      \end{enumerate}
    }

    \minor{\item[14.] Suppose the eigenvector matrix \tbm{S} has \(\bm{S}^T =
      \bm{S}^{-1}\). Show that \(\bm{A} = \bm{S\Lambda S}^{-1}\) is symmetric
      and has orthogonal eigenvectors.
    }
    \begin{align*}
      \bm{A} &= \bm{S\Lambda S}^{-1}  \\
      \bm{A}^T  &= (\bm{S\Lambda S}^{-1})^T   \\
      \bm{A}^T  &= (\bm{S}^{-1})^T \Lambda^T \bm{S}^T     \\
      \bm{A}^T  &= \bm{S} \Lambda \bm{S}^T = \bm{S} \Lambda \bm{S}^{-1}  && \bm{S}^T = \bm{S}^{-1}\\
                &\then \bm{A} = \bm{A}^T && ~\text{\tbm{A} is symmetric}~ \\\\
      \bm{S}^T \bm{S} &= \bm{S}^{-1}\bm{S} = \bm{I} && ~\text{\tbm{S} is orthonormal}~
    \end{align*}


    \minor{\item[39.] Suppose \(\bm{Ax} = \lambda \bm{x}\). If \(\lambda = 0,\)
      then \(x\) is in the nullspace. If \(\lambda \neq 0\), then \(x\) is in
      the column space. Those spaces have dimensions \((n-r)+r = n\). So why
      doesn't every square matrix have \(n\) linearly independent eigenvectors?
    }
    \begin{itemize}
      \item Because the dimensions of the \textit{row space} and null space
        must equal \(n\), the \textit{column space} and null space can overlap.
      \item This means that not every independent vector is an independent
        eigenvector. Non-distinct eigenvalues could map to the same
        eigenvector, but still yield independent columns in the original
        matrix.
    \end{itemize}

  \end{enumerate}
\end{itemize}

\section{5.5 Complex Matrices}
\begin{itemize}
  \item []

  \subsection{Problems 2, 12, 50}
  \begin{enumerate}
    \minor{\item[2.] What can you say about
      \begin{enumerate}
        \item the sum of a complex number and its conjugate?
        \begin{itemize}\color{foreground}
          \item is \(2x\) the real component; the imaginary component cancels out.
        \end{itemize}

        \item the conjugate of a number on the unit circle?
        \begin{itemize}\color{foreground}
          \ite also on the unit circle, but with an inverted handed angle
          (reflected about the real axis).
        \end{itemize}

        \item the product of two numbers on the unit circle?
        \begin{itemize}\color{foreground}
          \item is also on the unit circle, with real axis equal to the sum of
            angels of the factors of the product.
        \end{itemize}

        \item the sum of two numbers on the unit circle?
        \begin{itemize}\color{foreground}
          \item not on the unit circle, but lies on a circle that is less on or
            inside another circle with the radius of 2.
        \end{itemize}

      \end{enumerate}
    }

    \minor{\item[12.] Give a reason if true or a counterexample if false:
      \begin{enumerate}
        \item If \tbm{A} is Hermitian, then \(\bm{A} + \i \bm{I}\) is invertible.
        \begin{itemize}\color{foreground}
          \item \true{true}; \(\bm{A} + \i\bm{I}\) is guaranteed to have all
            non-zero complex entries along the diagonal, making it nonsingular.
        \end{itemize}

        \item If \tbm{Q} is orthogonal, then \(\bm{Q} +\frac{1}{2} \bm{I}\) is
          invertible.
        \begin{itemize}\color{foreground}
          \item \true{true}; \(\bm{Q}\bm{Q}^{-1} = \bm{Q}^{-1}\bm{Q} = \bm{I}\) remains true.
        \end{itemize}

        \item If \tbm{A} is real, then \(\bm{A} + \i \bm{I}\) is invertible.
        \begin{itemize}\color{foreground}
          \item \false{false}; counterexample:
            \[%%%%%%%%%%
            \bm{A} =
            \begin{bmatrix}
              0 & 1 \\
              -1 & 0
            \end{bmatrix} \then
            \bm{A} + \i \bm{I} =
            \begin{bmatrix}
              \i & 1 \\
              -1 & \i
            \end{bmatrix}
            \]%%%%%%%%%%
            if \(\i = 0\), then this matrix would remain singular.
        \end{itemize}

      \end{enumerate}
    }

    \minor{\item[50.] A matrix with the orthonormal eigenvectors has the form
      \(\bm{A} = \bm{U\Lambda U}^{-1} = \bm{U\Lambda U}^{\t}\). Prove that
      \(\bm{AA}^{\t} = \bm{A}^{\t} \bm{A}\). These are exactly the normal
      matrices.
    }
    \begin{align*}
      \bm{AA}^{\t} &= \bm{U\Lambda U}^{\t} (\bm{U\Lambda U}^{\t})^{\t}  \\
                   &= \bm{U\Lambda U}^{\t} \bm{U\Lambda^{\t}  U}^{\t}  \\
                   &= \bm{U\Lambda} \bm{\Lambda^{\t}  U}^{\t}
      = \aset{\bm{U\Lambda}^{\t}  \bm{\Lambda  U}^{\t}}
      && \bm{\Lambda\Lambda}^{\t} = \bm{\Lambda^{\t} \Lambda}\\\\
      \bm{A^{\t} A} &= (\bm{U\Lambda U}^{\t})^{\t}  \bm{U\Lambda U}^{\t}  \\
                    &= \bm{U \Lambda^{\t}  U}^{\t} \bm{U\Lambda  U}^{\t}  \\
                    &= \aset{\bm{U\Lambda}^{\t}  \bm{\Lambda  U}^{\t}}
                    &&\then \bm{AA}^{\t} = \bm{A^{\t} A}
    \end{align*}

  \end{enumerate}
\end{itemize}

\section{5.6 Similarity Transformations}
\begin{itemize}
  \item []

  \subsection{Problems 17, 21, 42}
  \begin{enumerate}
    \minor{\item[17.] Prove that every unitary matrix \tbm{A} is
      diagonalizable, in two steps:
      \begin{enumerate}
        \item If \tbm{A} is unitary, and \tbm{U} is too, then so it \(\bm{T} =
          \bm{U}^{-1}\bm{AU}.\)

        \item An upper triangular \(\bm{T}\) that is unitary must be diagonal.
          Thus, \(\bm{T}=\bm{A}\).

      \end{enumerate}

      Any unitary matrix \tbm{A} (distinct eigenvalues or not) has a complete set of orthonormal eigenvectors. All eigenvalues satisfy \(|\lambda | = 1\).

    }

    \minor{\item[21.] Prove that a matrix with orthonormal eigenvectors must be
      normal, as claimed: if \(\bm{U}^{-1} \bm{NU} = \bm{A}\) or \(\bm{N} =
      \bm{U\Lambda U}^{\t}\), then \(\bm{NN}^{\t} = \bm{N}^{\t}\bm{N}.\)
    }

    \minor{\item[42.] Prove that \tbm{AB} has the same eigenvalues as \tbm{BA}.
    }

  \end{enumerate}
\end{itemize}
